---
title: "ECO2460 - Growth Rates"
author: "Kristine Villaluna"
date: "18/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

## Exercise 3: Application - Prediction of Debt Crisis Events

The variable descriptions are as follows: 

* grossdebt2gdpweo: Gross debt (%GDP) 
* efn2gdp: external financing need EFN (% GDP) 
* externdebt2gdp: external debt (%GDP) 
* gfn2gdp: gross financing need GFN (% GDP) 
* interExtdeb2gdp: Interest on external debt (% GDP) 
* CA2gdp: Current account (% GDP) 
* debtserv2expor: Debt service/Exports (%) 
* credit2privat: Credit to private sector (% GDP) 
* FCD2GDweo : Foreign currency denominated debt (% public debt) 
* IntExp2rev: Interest Expenditure/Government revenue (%) 
* debtserv2res: Debt services/Reserves (%) 
* stdebt2extdebt: Short term debt /External debt (%) 

### 8. Calculating Growth Rates

```{r}
library(readxl)
library(here)
library(janitor)
library(data.table)
library(dplyr)
```

```{r}
crisis <- read.csv(here("crisis_ohe_imputed.csv"))
crisis <- clean_names(crisis)

crisis2 <- read_excel(here("crisisdata.xls"))
crisis2 <- clean_names(crisis2)
crisis2 <- crisis2 %>% select(country)

crisis3 <- cbind(crisis,crisis2)
```

```{r}
# calculate growth rate for each variable
growth <- as.data.table(crisis3)[, lapply(.SD, function(x)x/shift(x) - 1), .SDcols = 3:14, by=country]

# rename new growth variables
colnames(growth)<-paste(colnames(growth),"growth",sep="_")

# combine data frames
crisis_final <- cbind(crisis3,growth)

# remove country growth
crisis_final <- crisis_final %>% dplyr::select(-country_growth) %>% select(-country)
```

```{r}
# export new csv
write.csv(crisis_final,"crisis_growth.csv", row.names = FALSE)
```


